---
title: "Data Structures"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Vectors 

```{r}
(grand_slam_win <- c(9, 15, 5, 12, 18))
```
```{r}
a = c(1,2,3)
b = c(4,5,6)
(d = c(a,b,a))
(d1 = c("a", "b", "a")) 
```

##### On Types 

Types can be double = real number or character (and others)
```{r}
typeof(d)
typeof(d1)
typeof(grand_slam_win)
```

When creating a vector with different types of element there an order to transform some of them in order to make them all the same (see book sec 4.1.2 coercion)

```{r}
(d2 = c(a, d1))
typeof(d2)
```

##### Subsetting (subset of vectors)

```{r}
(a = c(321, 14, 1134, 465, 643, 6343, 64))
a[3]
a[c(5, 1, 3)]
a[-3]
a[-c(1, 2)]
```

We cannot do this a[c(1, 2, -3)], but we can do

```{r}
a[c(1, 2, 4)] [-3]
a[c(1, 2, 4)] [-4]
```


Another way of subsetting

```{r}
a[c(F, T, F, T, F, F, F)]
a[a > 600]
a > 600
```


##### Attributes 

```{r}
(grand_slam_win <- c("Andy Murray" = 9, "Rafael Nadal" = 15, 
                   "Stan Wawrinka" = 5, "Novak Djokovic" = 12, 
                   "Roger Federer" = 18))
```

Same commande
```{r}
grand_slam_win[4]
grand_slam_win["Novak Djokovic"]
```

```{r}
length(grand_slam_win)
sort(grand_slam_win)
order(grand_slam_win)
```

Sort can be done like that :)

```{r}
grand_slam_win[order(grand_slam_win)]
```

##### Sequences

The last one is taking equidistant value between 3 and 7.4

```{r}
1:6
6:1
seq(from = 3, to = 7.4, length.out = 10)
```

Say you want to have the date of three months ago

1. create the vector today
```{r}
(today = Sys.Date())
```

2. extract the date of three months ago
```{r}
(three_months_ago = seq(today, length = 2, by = "-3 months" )[2])
```

##### Exemple 

```{r}
library(quantmod)
getSymbols("AAPL", from = three_months_ago, to = today)
candleChart(AAPL, theme = 'white', type = 'candles')
```

```{r}
AAPL
ClCl(AAPL)
mean(ClCl(AAPL))
mean(na.omit(ClCl(AAPL)))
```


##### Matrices

We can create matrices with vectors
```{r}
players <- c("Andy Murray", "Rafael Nadal", "Stan Wawrinka", 
             "Novak Djokovic", "Roger Federer")
grand_slam_win <- c(9, 15, 5, 12, 18)
win_percentage <- c(78.07, 82.48, 63.96, 82.77, 81.80)
(mat <- cbind(grand_slam_win, win_percentage))
```

We can give other names to column or rows

```{r}
rownames(mat) <- players
colnames(mat) <- c("GS win", "Win rate")
mat
```

Here's another way (the [[2]] correspond to the column, [[1]] would be the row)

```{r}
dimnames(mat)[[2]] <- c("GS wins", "Win rate")
mat
```

Matrice transpose and matrice product

```{r}
t(mat)
mat%*%t(mat)
```


##### Exemple

```{r}
# Load quantmod
library(quantmod)

# Download data
today <- Sys.Date()
five_years_ago <- seq(today, length = 2, by = "-5 year")[2]
getSymbols("AAPL", from = five_years_ago, to = today)
getSymbols("NFLX", from = five_years_ago, to = today)

# Compute returns
Ra <- na.omit(ClCl(AAPL))
Rn <- na.omit(ClCl(NFLX)) 

# Estimation of mu and Sigma
Sigma <- cov(cbind(Ra, Rn))
mu <- c(mean(Ra), mean(Rn))
Sigma
mu

# Compute omega^*
omega_star <- (Sigma[2, 2] - Sigma[1, 2])/(Sigma[1, 1] + Sigma[2, 2] - 2*Sigma[1, 2])
omega_star


# Compute investment expected value and variance
mu_investment <- omega_star*mu[1] + (1 - omega_star)*mu[2]
var_investment <- omega_star^2*Sigma[1,1] + (1 - omega_star)^2*Sigma[2,2] + 
  2*omega_star*(1 - omega_star)*Sigma[1,2]


investment_summary <- matrix(NA, 2, 3)
dimnames(investment_summary)[[1]] <- c("Expected value", "Variance")
dimnames(investment_summary)[[2]] <- c("Apple", "Netflix", "Investment")
investment_summary[1, ] <- c(mu, mu_investment)
investment_summary[2, ] <- c(diag(Sigma), var_investment)
knitr::kable(investment_summary)
```

##### List

A list is a heterogenious object

```{r}
# List elements
num_vec <- c(188, 140)
char_vec <- c("Height", "Weight", "Length")
logic_vec <- rep(TRUE, 8)
my_mat <- matrix(0, nrow = 5, ncol = 5)

# List initialization 
(my_list <- list(num_vec, char_vec, logic_vec, my_mat))
```

Subsetting list

1. give name to the elements of the list
```{r}
(my_list <- list(number = num_vec, character = char_vec, 
                 logic = logic_vec, matrix = my_mat))
```

2. subsetting (two ways)
```{r}
my_list$character
my_list[[2]]
```

The good thing with List is that the elements don't need to be of the same length.

##### Data Frames

If one the contrary the elements of a list have the same length, the we can use a data frame instead

```{r}
### Creation

players <- c("Andy Murray", "Rafael Nadal", "Stan Wawrinka", 
             "Novak Djokovic", "Roger Federer")

grand_slam_win <- c(9, 15, 5, 12, 18)

date_of_birth <- c("15 May 1987", "3 June 1986", "28 March 1985", 
                  "22 May 1987", "8 August 1981")

country <- c("Great Britain", "Spain", "Switzerland", 
            "Serbia", "Switzerland")
ATP_ranking <- c(1, 2, 3, 4, 5)

prize_money <- c(60449649, 85920132, 30577981, 109447408, 104445185)

tennis <- data.frame(date_of_birth, grand_slam_win, country, 
                    ATP_ranking, prize_money)
tennis
```

If we want to give names to the rows, we do 
```{r}
# Attenttion: dimnames(tennis) is a list, whereas tennis is a data frame
dimnames(tennis)[[1]] <- players
tennis
```

Check the type
```{r}
is.list(2)
is.list(tennis)
is.data.frame(tennis)
```


##### Graphs

We are going to redo the graph of the less risky investement strategy 
```{r}
# load the x-value and y-value. 
# We can use the following table investement_summary
# etc...
```

##### Code

We start with logical operator between number of 1-dim object
```{r}
a = 3
b = 4
a < b
!(a > b) || b < 0
!(a > b) && b < 0
```

If we have vectors
```{r}
a = 1:12
a > 10
```
it can be problematic cause here only the first value of the vecor is checked.
```{r}
a > 5 && a < 10
```
To do it with vectors we use & instead of &&
```{r}
a > 5 & a < 10
a[a > 5 & a < 10]
sum(a > 5 & a < 10)
```

Code if a number is even or odd
```{r}
a = 4
if(a %% 2 == 0){
  print("even")
}else{
  print("odd")
}
```

Or we can do and use the paste function since a and "even" are not of the same type
```{r}
a = 4
if(a %% 2 == 0){
  print(paste(a,"is even"))
}else{
  print(paste(a,"is odd"))
}
```

Using for and if
```{r}
a = 2
for(i in 1:6){
  if(a > 5){
    a = 2
  }
  a = a+1
}
a
```

We are going to make the sum of the elements of a vector in two ways

1. simple
```{r}
a = 1:7
sum(a)
```

2. tricky
```{r}
a = 1:6
mysum = 0
for(i in 1:length(a)){
  mysum = mysum + a[i]
}
mysum
```

